### 2019-02-14 2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/main.tex
Okay je réorganise et réécris pas mal pour ce chapitre 4.

In this Section, we present a demonstration showed at the International Conference on Telecommunication (ICT) in June $2018$, implementing a proof-of-concept of the model previously presented in Section~\ref{sec:4:firstModel}.

We implement an IoT network the following way: one gateway, one or several intelligent (\emph{i.e.}, learning) objects, embedding the proposed solution,
and a traffic generator that emulates radio interference from many other objects.
Intelligent objects communicate with the gateway with a wireless ALOHA-based protocol, which does not require any specific overhead for the learning.
%
We model the network access as a discrete sequential decision making problem, and using the framework and algorithms from MAB learning, we show that intelligent objects can improve their access to the network by using low complexity and decentralized algorithms, such as \UCB{} and Thompson Sampling.
%
This solution could be added in a straightforward and costless manner in LoRaWAN networks, just by adding this feature in some or all the devices, without any modification on the network side.

\TODOL{Based on the publication, ``MALIN: Multi-Arm bandit Learning for Iot Networks with GRC: A TestBed Implementation and Demonstration that Learning Helps'', demo at ICT, and the companion paper ``GNU Radio Implementation of MALIN: "Multi-Armed bandits Learning for Internet-of-things Network"'', see https://hal.inria.fr/hal-02006825}

\TODOL{TODO:

- Clean up notations, be sure it is OK with previous Sections

- Be sure to not repeat any motivations / maths / explanations given before

- Be sure to give enough details about the conclusions etc
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Motivation for this demonstration}

% The monitoring of large scale systems, such as smart grids and smart cities, requires the development of networks dedicated to Internet-of-Things (IoT) applications.
% For instance, Low Power Wide Area Networks (LPWAN) \cite{Raza17}, as LoRaWAN or SigFox, are nowadays deployed in unlicensed bands to handle a large number of objects transmitting a few packets per day or week.
% In order to reduce the energy consumption of end-devices, these networks rely on pure ALOHA-based Medium Access (MAC) protocols.

% One of the challenges in the design of MAC solutions for the IoT is to design solutions which improve the performance of the network and reduce the Packet Loss Ratio (PLR), without reducing the end-devices battery life.
% In particular, many IoT standards operate in unlicensed bands, that is why we have to find solutions that do not increase the PLR due to the interference caused by other standards and networks which share the same band, without coordination.
% As this interfering traffic is generated by different standards and networks, it cannot be controlled, and it is not evenly distributed in the different channels.

Multi-Armed Bandit (MAB) algorithms have been recently proposed as a solution to improve the performance of IoT networks and in particular in LPWAN \cite{Bonnefoi18,Azari18}.
In this paper, we describe the way we implemented a demo where we evaluate MAB algorithms, used in combination with a pure ALOHA-based protocol (such as the ones employed in LPWAN).
%
%\todo[inline]{Contributions et interets de la demo}
% MAB learning has already been proposed in Cognitive Radio (CR) \cite{Haykin}, and in particular, for sensing-based Dynamic Spectrum Access (DSA) in licensed bands \cite{Jouini}.
% Recently, TS and UCB{} algorithms have been used for improving the spectrum access in (unlicensed) WiFi networks \cite{Toldov}, and the UCB{} algorithm was used in a unlicensed and frequency- and time-slotted IoT network \cite{Bonnefoi}.
% Many recent works show that MAB algorithms work well for real-world radio signal.
% However, even with only one dynamic user using the learning algorithm, the background traffic or the traffic of the other objects is never really stationary or \emph{i.i.d.} (independent and identically distributed).
% In recent works like \cite{Bonnefoi}, several objects are using bandit algorithms, and the assumptions made by the stochastic bandit algorithms are not satisfied: as several agents learn simultaneously, their behavior is neither stationary nor \emph{i.i.d.}.
% As far as we know, we provide the first study to confirm robustness of the use of stochastic bandit algorithms for decision making in IoT networks with a large number of intelligent objects in the network, which makes the environment not stationary at all, violating the hypothesis required for mathematical proofs of bandit algorithms convergence and efficiency.
%
This demonstration is the first implementation which aims at assessing the potential gain of MAB learning algorithms in IoT scenarios.
% % even when the number of intelligent objects in the network increases, and the stochastic hypothesis is more and more questionable.
% To do that, we suppose an IoT network made of two types of objects: static objects that use only one channel (fixed in time), and dynamic objects that can choose the channel for each of their transmissions. Static objects form an interfering traffic, which could have been generated by objects using other standards as well.
% We first evaluate the probability of collision if dynamic objects randomly select channels (naive approach), and if a centralized controller optimally distribute them in channels (ideal approach).
% Then, these reference scenarios allow to evaluate the performance of UCB and TS algorithms in a decentralized network, in terms of successful communication rate, as it reflects the network efficiency.
% We show that these algorithms have near-optimal performance, even when the proportion of end-objects increases and the interfering traffic from other objects becomes less and less stochastic.
% \todo[inline]{(Remi) Si on parle de non stationarity, il faut donner un resultat a deux utilisateurs.}
Following our recent work presented in the previous Section~\ref{sec:4:firstModel}, we propose to model this problem as Non-Stationary\footnote{~Note that non-stationarity only comes from the presence of more than one dynamic object, as the background traffic is assumed independent and identically distributed \emph{i.i.d.}.} MAB.
We suggest to use low-cost algorithms, focusing on two well-known algorithms: a frequentist one (\UCB{}) and a Bayesian one (TS).
We consider the Upper-Confidence Bound (\UCB{}) \cite{Auer}, and the Thompson Sampling (TS) algorithms \cite{Thompson33}. Both algorithms have already been applied with success in the context of wireless decision making, both empirically for Opportunistic Spectrum Access \cite{Jouini},
and more recently for multi-users Cognitive Radio problems \cite{BessonALT18} with a more theoretical approach.

We use a TestBed designed in 2017 by our team SCEE \cite{Bodinier17}, containing different USRP boards \cite{USRPDocumentation}, controlled by a single laptop using GNU Radio \cite{GNURadioDocumentation},
and where the intelligence of each object corresponds to a learning algorithm, implemented as a GNU Radio block \cite{GNURadioCompanionDocumentation} and written in Python or \texttt{C++}.

@@

The rest of this paper is organized as follows. The system model is introduced in Section 2. In Section 3, we describe more formally both the \UCB{} and the TS algorithms. Our implementation is presented in Section 4, and results are given in Section 5.
% and Section 6 concludes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{System model - FIXME merge with previous section!}

We consider the system model presented in Figure~\ref{fig:42:system_model1}, where a set of object sends uplink packets to the network gateway, in the $433.5\;\mathrm{MHz}$ ISM band.
The communication between IoT devices and this gateway is done through a simple pure ALOHA-based protocol where devices transmit uplink packets of fixed duration whenever they want.
%
The devices can transmit their packets in $K\geq 1$ channels (\emph{e.g.}, $K=4$). In the case where the gateway receives an uplink in one channel, it transmits an acknowledgement to the end-device in the same channel, after a fixed delay (of $1$ s).

These communications operate in unlicensed ISM bands and, consequently, suffer from interference generated by uncoordinated neighboring networks. This interfering traffic is uncontrolled, and can be unevenly distributed over the $K$ different channels.

%To simulate networks designed for the Internet of Things (IoT), we consider a protocol with no sensing, no repetition of uplink messages, and where the gateway is in charge of sending back an acknowledgement, after some fixed-time delay (typically $1\;\mathrm{s}$), to any object who succeeded in sending successfully an uplink packet.
%
%By considering a small number of wireless channels (\emph{e.g.}, $K=8$) and one PHY layer configuration (\emph{i.e.}, modulation, waveform, etc), and in case of a non-uniform traffic in the different channels,
%the object can improve their usage of the network if they are able to \emph{learn} on the fly the best channels to use (\emph{i.e.}, the most vacant one).

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/system_model1.eps}
    \caption{In our system model, some dynamic devices transmit packets to a gateway and suffer from the interference generated by neighboring networks.}
    \label{fig:42:system_model1}
\end{figure}

We consider the network from the point of view of one end-user. Every times the end-user has to communicate with the gateway,
it has to choose one channel (at each transmission $t \geq 1, t \in \mathbb{N}$), denoted as $C(t) = k \in\{1,\dots,K\}$.
Then, the end-users starts waiting in this channel $C(t)$ for an acknowledgement sent by the gateway.
Before sending another message (\emph{i.e.}, at time $t+1$), the end-user knows if it received or not this \emph{ACK} message.
%
For this reason, selecting channel (or arm) $k$ at time $t$ yields a (random) feedback, called a \emph{reward}, $r_k(t) \in \{0,1\}$, being $0$ if no \emph{ACK} was received before the next message, or $1$ if \emph{ACK} was successfully received.
The goal of the end-user is to minimize its packet loss ratio, or equivalently, it is to maximize its cumulative reward,
$r_{1 \dots T} := \sum_{t = 1}^T r_{C(t)}(t),$
as it is usually done in MAB problems \cite{Thompson33,Robbins52,LaiRobbins85}.

This problem is a special case of the so-called ``stochastic'' MAB, where the sequence of rewards drawn from a given arm $k$ is assumed to be  \emph{i.i.d.}, under some distribution $\nu_k$, that has a mean $\mu_k$. Several types of reward distributions have been considered in the literature, for example distributions that belong to a one-dimensional exponential family (\emph{e.g.}, Gaussian, Exponential, Poisson or Bernoulli distributions).

Rewards are binary in our model, and so we consider only Bernoulli distributions, in which $r_k(t) \sim \mathrm{Bern}(\mu_k)$, that is, $r_k(t) \in \{0,1\}$ and $\mathbb{P}(r_k(t) = 1) = \mu_k \in [0,1]$.
Contrary to many previous work done in the CR field (\emph{e.g.}, Opportunistic Spectrum Access),
the reward $r_k(t)$ does \emph{not} come from a sensing phase before sending the $t$-th message, as it would do for any ``listen-before-talk'' model. Rewards come from receiving an acknowledgement from the gateway, between the $t$-th and $t+1$-th messages.

The problem parameters $\mu_1,\dots,\mu_K$ are of course unknown to the end-users, so to maximize its cumulated reward, it must learn the distributions of the channels, in order to be able to progressively focus on the best arm (\emph{i.e.}, the arm with largest mean).
%
This requires to tackle the so-called \emph{exploration-exploitation dilemma}: a player has to try all arms a sufficient number of times to get a robust estimate of their qualities, while not selecting the worst arms too many times.

% In a Cognitive Radio application, arms model the \emph{channels}, and players are the \emph{dynamic end-devices}.
% For example in the classical OSA setting with sensing \cite{Jouini}, a single dynamic device (a player) sequentially tries to access channels (the arms), and collects a reward of 1 if the channel is available %(as there is no primary user, we can assume the transmission to be successful),
% and 0 otherwise. So rewards represent the \emph{availability} of channels, and
% the parameter $\mu_k$ represents the mean availability of channel $k$.

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsection{MAB Algorithms}

Before discussing the relevance of a MAB model for our IoT application, we refer to Section~\ref{sec:2:famousMABalgorithms} which presented two bandit algorithms, \UCB{} and Thompson Sampling,
which are both known to be efficient for stationary \emph{i.i.d.} rewards and are shown to be useful in our setting (in Sec.~\ref{sec:42:results}).

% \paragraph{The \UCB{} Algorithm}\label{sub:42:UCB}

% A naive approach could be to use an empirical mean estimator of the rewards for each channel, and select the channel with highest estimated mean at each time;
% but this greedy approach is known to fail dramatically \cite{LaiRobbins85}. Indeed, with this policy, the selection of arms is highly dependent on the first draws:
% if the first transmission in one channel fails and the first one on other channels succeed, the end-user will \emph{never} use the first channel again, even it is the best one (\emph{i.e.}, the most available, in average).

% Rather than relying on the empirical mean reward, Upper Confidence Bounds algorithms
% instead use a \emph{confidence interval} on the unknown mean $\mu_k$ of each arm,
% which can be viewed as adding a ``bonus'' exploration to the empirical mean.
% % by adding a confidence term to the empirical mean.
% They follow the ``\emph{optimism-in-face-of-uncertainty}'' principle: at each step, they play according to the best model,
% as the statistically best possible arm (\emph{i.e.}, the highest upper confidence bound) is selected.
% % the arm with the highest upper confidence bound is expected to be the true best arm, and is played.

% More formally, for one end-user, let $N_k(t) = \sum_{\tau=1}^t \mathbbm{1}(C(\tau) = k)$ be the number of times channel $k$ was selected up-to time $t \geq 1$.
% The empirical mean estimator of channel $k$ is defined as the mean reward obtained by selecting it up to time $t$, $\widehat{\mu_k}(t) = \left(1 / N_k(t)\right) \sum_{\tau=1}^t r_k(\tau) \mathbbm{1}(C(\tau) = k) $.
% For \UCB, the \emph{confidence} term is $B_k(t) = \sqrt{\alpha \log(t) / N_k(t)}$,
% giving the upper confidence bound $U_k(t) = \widehat{\mu_k}(t) + B_k(t)$, which is used by the end-user to decide the channel for communicating at time step $t+1$: $C(t+1) = \arg\max_{1\leq k \leq K} U_k(t)$.
% \UCB{} is called an \emph{index policy}.

The \UCB{} algorithm uses a parameter $\alpha > 0$, originally $\alpha$ was set to $2$ \cite{Auer}, but empirically $\alpha = 1/2$ is known to work better (uniformly across problems), and $\alpha > 1/2$ is advised by the theory \cite{bubeck2012regret}.
% This algorithm is simple to implement and use in practice, even on embedded micro-processors with limited computation and memory capabilities.
%
% \todo[inline]{Maybe in the implementation of UCB the $t$ $\log(t)$ should be the total amount of time the user tried to communicate? It is already the case, but we should say so.}
In our model, every dynamic end-user implements its own \UCB{} algorithm, \emph{independently}.
For one end-user, the time $t$ is the total number of sent messages from the beginning, as rewards are only obtained after a transmission.

% \paragraph{Thompson Sampling}

% Thompson Sampling \cite{Thompson33} was introduced early on, in $1933$ as the very first bandit algorithm, in the context of clinical trials (in which each arm models the efficacy of one treatment across patients). Given a prior distribution on the mean of each arm, the algorithm selects the next arm to draw based on samples from the \emph{conjugated} posterior distribution, which for Bernoulli rewards is a Beta distribution.

% A Beta prior $\mathrm{Beta}(a_k(0)=1,b_k(0)=1)$ (initially uniform) is assumed on $\mu_k \in [0, 1]$, and at time $t$ the posterior is $\mathrm{Beta}(a_k(t),b_k(t))$.
% After every channel selection, the posterior is updated to have $a_k(t)$ and $b_k(t)$ counting the number of successful and failed transmissions made on channel $k$.
% So if the \emph{ACK} message is received, $a_k(t+1) = a_k(t) + 1$, and $b_k(t+1) = b_k(t)$, otherwise $a_k(t+1) = a_k(t)$, and $b_k(t+1) = b_k(t) + 1$.
% Then, the decision is done by \emph{sampling} an \emph{index} for each arm, at each time step $t$, from the arm posteriors: $X_k(t) \sim \mathrm{Beta}(a_k(t), b_k(t))$, and the chosen channel is simply the channel $C(t+1)$ with highest index $X_k(t)$. For this reason, Thompson Sampling is called a \emph{randomized index policy}.

The TS algorithm, although being simple and easy to implement, is known to perform well for stochastic problems, for which it was proven to be asymptotically optimal \cite{AgrawalGoyal11,Kaufmann12}.
It is known to be empirically efficient, and for these reasons it has been used successfully in various applications, including on problems from Cognitive Radio \cite{Toldov,Mitton}, and also in previous work on decentralized IoT-like networks \cite{Darak16}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{GNU Radio implementation}

In this section, we present our implementation of MAB algorithms in our model of IoT networks.
We first describe the simplified physical layer of this demo,
then we present our GNU Radio implementation.

@@

In this paper, we implement a PHY/MAC layers solution in order to improve the performance of IoT communications in unlicensed bands. We could have used any physical layer and any ALOHA-based protocol.
We choose to implement our own physical layer and protocol, for both clarity and conciseness.

@@

\paragraph{Equipment}

% Boards and versions
We use USRP N210 boards \cite{USRPDocumentation}, from Ettus Research (National Instrument),
with version $4$ of their FPGA system and version $5.1$ of the RBX system.
As illustrated in Figure~\ref{fig:42:our_demo}, our implementation is composed of at least $3$ USRP.
The gateway, a USRP which emulates the interfering traffic, and at least one dynamic device.
% Cables and laptop

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.70\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/our-demo.eps}
    \caption{Schematic of our implementation that presents the role of each USRP card.}
    \label{fig:42:our_demo}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/SCEE_TestBed1.jpg}
    \vspace*{20pt}
    \includegraphics[width=0.65\linewidth]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/SCEE_TestBed2.jpg}
    \caption{Photos of the SCEE testbed.}
    \label{fig:42:photosSCEETestBed}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.80\textwidth]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/UI.eps}
    \caption{User interface of our demonstration.}
    \label{fig:42:UI}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Results and discussions}\label{sec:42:results}

@@

Similar results are obtained for overlapping channels, when dynamic devices are learning in the presence of multiple devices, all using the same learning algorithm.
Empirical results confirm the simulations presented in our paper \cite[Fig.3]{Bonnefoi17}.
Such results are very encouraging, and illustrate well the various strong possibilities of MAB learning applied to IoT networks.

\begin{figure}[!t]
	\centering
    \includegraphics[height=8.0cm]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
    \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) suffice for the two learning objects to reach a successful communication rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) object, which stays around $40\%$ of success. Similar gains of performance were obtained in many different scenarios.}
    \label{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}
\end{figure}

% \begin{figure}[!t]
% 	\centering
%     \begin{subfigure}[b]{0.46\textwidth}
%         \includegraphics[height=7cm]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/Object2_UniformAccess__UI.png}
%         %\caption{Object using uniform access.}
%         \label{fig:42:comparing_UniformAccess_and_UCB__1}
%     \end{subfigure}
%     ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
%       %(or a blank line to force the subfigure onto a new line)
%     \begin{subfigure}[b]{0.46\textwidth}
%         \includegraphics[height=7cm]{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/Object1_UCB__UI.png}
%         %\caption{Object using UCB.}
%         \label{fig:42:comparing_UniformAccess_and_UCB__2}
%     \end{subfigure}
% \caption{Comparing the success rate of an object using uniform access (left, $41\%$ of success here) and an object using the UCB algorithm (right, already $60\%$ of success). We see the learning object being much more present in the channels $\#4$ and $\#3$ (\textcolor{darkblue}{blue chart}, bottom right corner), as these channels are the less perturbed by the random interfering traffic, in this example.}
% \label{fig:42:comparing_UniformAccess_and_UCB}
% \end{figure}

We presented in this Section a demonstration, showed in $2018$ at the ICT conference,
by specifying the system model and explaining the two MAB algorithms used in practice.
We gave all the necessary details on both the PHY and the MAC layer, as well as details on the User Interface developed for the demo.
Results obtained in practice were discussed, to highlight the interest of using learning algorithms for radio online optimization problem, and especially multi-armed bandit learning algorithms.
%
By using such low-cost algorithms, we demonstrated empirically that a dynamically reconfigurable object can learn on its own to favor a certain channel, if the environment traffic is not uniform between the $K$ different channels.

Possible future extensions of this work include:
considering more dynamic objects (\emph{e.g.}, $100$),
implementing a real-world IoT communication protocol (like the LoRaWAN standard),
and studying the interference in case of other gateways located nearby.
%
% Pub pour notre autre article
We are also interested in studying the possible gain of using a learning step when the transmission model follows ALOHA-like retransmissions, and this is presented in the next Section~\ref{sec:4:retransmissions}.

\paragraph{Availability of data and materials}
%
The source code of our demonstration is fully available online, open-sourced under GPLv3 license, at\\
\href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc/}}.
It contains both the GNU Radio Companion flowcharts and blocks, with ready-to-use \texttt{Makefiles} to easily compile, install and launch the demonstration.

A $6$-minute \textbf{video} showing our demonstration is at \texttt{\url{youtu.be/HospLNQhcMk}}.
It shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Fig.~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
It shows the connections between the USRP boards, the Octoclock, and the master laptop.

@@

% ---------- Bibliography ----------
%\begin{small} % XXX small if needed?
\bibliographystyle{ieeetr}
\bibliography{biblio_RIoT}
%\end{small} % XXX small if needed?

>>>>

\graphicspath{{2-Chapters/4-Chapter/IEEE_WCNC_2019__DemoICT.git/pictures/}}

In this Section, we present a demonstration showed at the International Conference on Telecommunication (ICT) in June $2018$ \cite{Besson2018ICT,Besson2019WCNC}, implementing a proof-of-concept of the model previously presented in Section~\ref{sec:4:firstModel}.

We implement an IoT network the following way: one gateway, one or several intelligent (\emph{i.e.}, learning) objects, embedding the proposed solution,
and a traffic generator that emulates radio interference from many other objects.
Intelligent objects communicate with the gateway with a wireless ALOHA-based protocol, which does not require any specific overhead for the learning.
%
Similarly to the previous section, network access is modelled as a discrete sequential decision making problem, and using the framework and algorithms from MAB learning, we show that intelligent objects can improve their access to the network by using low complexity and decentralized algorithms, such as \UCB{} and Thompson Sampling.
%
This solution could be added in a straightforward and costless manner in LoRaWAN networks, just by adding this feature in some or all the devices, without any modification on the network side.

% \TODOL{TODO:

% - Clean up notations, be sure it is OK with previous Sections

% - Be sure to not repeat any motivations / maths / explanations given before

% - Be sure to give enough details about the conclusions etc
% }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Motivation for this demonstration}
\label{sub:42:motivation}

In this Section, we describe the way we implemented a demo where we evaluate MAB algorithms, used in combination with a pure ALOHA-based protocol, such as the ones employed in LPWAN.
This demonstration is the first implementation which aims at assessing the potential gain of MAB learning algorithms in IoT scenarios.
%
We use a TestBed designed in 2017 by our team SCEE \cite[Appendix~3]{Bodinier17}, containing different USRP boards \cite{USRPDocumentation}, controlled by a single laptop using GNU Radio \cite{GNURadioDocumentation},
and where the intelligence of each object corresponds to a learning algorithm, implemented as a GNU Radio block \cite{GNURadioCompanionDocumentation} and written in Python or \texttt{C++}.

@@

We refer to Section~\ref{sec:2:famousMABalgorithms} which presented two bandit algorithms, \UCB{} and Thompson Sampling,
which are both known to be efficient for stationary \emph{i.i.d.} rewards and are shown to be useful in our setting (in Section~\ref{sec:42:results}).
Then we discuss the relevance of a MAB model for our IoT application.

The rest of this Section is organized as follows.
% The system model is introduced in Section 2. In Section 3, we describe more formally both the \UCB{} and the TS algorithms.
Our implementation is presented in Section~\ref{sub:42:implementation}, with details about the user interface, and results are given in Section~\ref{sub:42:results}.
% and Section 6 concludes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{GNU Radio implementation}
\label{sub:42:implementation}

In this section, we present our implementation of MAB algorithms in the model of IoT networks presented in Section~\ref{sub:41:systemModel}.
We first describe the simplified physical layer of this demo, then we present our GNU Radio implementation.

@@

We implement a PHY/MAC layers solution in order to improve the performance of IoT communications in unlicensed bands. We could have used any physical layer and any ALOHA-based protocol.
We choose to implement our own physical layer and protocol, for both clarity and conciseness.

@@

\paragraph{Equipment}
% Boards and versions
We use USRP N210 boards \cite{USRPDocumentation}, from Ettus Research (National Instrument),
with version $4$ of their FPGA system and version $5.1$ of the RBX system.
As illustrated in Figure~\ref{fig:42:our_demo}, our implementation is composed of at least $3$ USRP.
The gateway, a USRP which emulates the interfering traffic, and at least one dynamic device.
% Cables and laptop
% Osef

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.70\linewidth]{our-demo.eps}
    \caption{Schematic of our implementation that presents the role of each USRP card.}
    \label{fig:42:our_demo}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.65\linewidth]{SCEE_TestBed1.jpg}
    \vspace*{20pt}
    \includegraphics[width=0.65\linewidth]{SCEE_TestBed2.jpg}
    \caption{Two pictures showing the SCEE testbed \cite[Appendix~3]{Bodinier17}.}
    \label{fig:42:photosSCEETestBed}
\end{figure}

@@

\begin{figure}[!t]
    \centering
    \includegraphics[width=0.80\textwidth]{UI.eps}
    \caption{User interface of our demonstration.}
    \label{fig:42:UI}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Numerical results}
\label{sec:42:results}

@@

Similar results are obtained for overlapping channels, when dynamic devices are learning in the presence of multiple devices, all using the same learning algorithm.
Empirical results confirm the simulations presented in Section~\ref{sec:4:firstModel} (see Figure~\ref{fig:41:perf_learning}).
%  \cite[Fig.3]{Bonnefoi17}.
Such results are very encouraging, and illustrate well the various strong possibilities of MAB learning applied to IoT networks.

\begin{figure}[!t]
	\centering
    \includegraphics[height=8.0cm]{plot_datafile_append_Uniform_vs_UCB_vs_TS.eps}
    \caption{Less than $400$ communication slots (\emph{i.e.}, less than $100$ trials in each channel) suffice for the two learning objects to reach a successful communication rate close to $80\%$, which is \textbf{twice as much} as the non-learning (uniform) object, which stays around $40\%$ of success. Similar gains of performance were obtained in many different scenarios.}
    \label{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS}
\end{figure}

% \begin{figure}[!t]
% 	\centering
%     \begin{subfigure}[b]{0.46\textwidth}
%         \includegraphics[height=7cm]{Object2_UniformAccess__UI.png}
%         %\caption{Object using uniform access.}
%         \label{fig:42:comparing_UniformAccess_and_UCB__1}
%     \end{subfigure}
%     ~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc.
%       %(or a blank line to force the subfigure onto a new line)
%     \begin{subfigure}[b]{0.46\textwidth}
%         \includegraphics[height=7cm]{Object1_UCB__UI.png}
%         %\caption{Object using UCB.}
%         \label{fig:42:comparing_UniformAccess_and_UCB__2}
%     \end{subfigure}
% \caption{Comparing the success rate of an object using uniform access (left, $41\%$ of success here) and an object using the UCB algorithm (right, already $60\%$ of success). We see the learning object being much more present in the channels $\#4$ and $\#3$ (\textcolor{darkblue}{blue chart}, bottom right corner), as these channels are the less perturbed by the random interfering traffic, in this example.}
% \label{fig:42:comparing_UniformAccess_and_UCB}
% \end{figure}

\paragraph{Availability of data and materials}
%
The source code of our demonstration is fully available online, open-sourced under GPLv3 license, at
\href{https://bitbucket.org/scee_ietr/malin-multi-arm-bandit-learning-for-iot-networks-with-grc}{\texttt{bitbucket.org/scee\_ietr/\\malin-multi-arm-bandit-learning-for-iot-networks-with-grc/}}.
It contains both the GNU Radio Companion flowcharts and blocks, with ready-to-use \texttt{Makefiles} to easily compile, install and launch the demonstration.

A $6$-minute \textbf{video} showing our demonstration is at \texttt{\url{youtu.be/HospLNQhcMk}}.
It shows examples of $3$ dynamic devices learning simultaneously, confirming the results of Fig.~\ref{fig:42:plot_datafile_append_Uniform_vs_UCB_vs_TS} for overlapping channels.
It especially shows the connections between the USRP boards, the Octoclock, and the master laptop.

@@
